\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-21 notes/2loopNLSM.tex
clarified cubic vs color-dual

In color-charged theories, both sides of the cut can be further
decomposed according to the color algebra.  For theories charged under
the adjoint of $SU(N)$ via $f^{abc}$-dressed amplitudes, it is
convenient to project onto a specific element of the Del
Duca-Dixon-Maltoni (DDM) color basis \cite{DixonMaltoni} by only
inserting color-ordered amplitudes, a la the right hand side of
\cref{eq:ddm-basis} in the product of trees, and restricting the set
of compatible diagrams to those whose color factors reduce to the
appropriate DDM element when summing over \emph{only the uncut color
  contractions}.  In this situation, the color becomes an overall
factor on the entire cut equality and thus can be ignored.  Cuts that
are organized in this manner are known as ``color-ordered'' cuts, in
the same sense as color-ordered tree amplitudes.  Since the DDM basis
(and dual Kleiss-Kuijf amplitude basis \cite{Kleiss:1988ne}) have
$(n-2)!$ elements, one generally needs to evaluate
$(n_1-2)! (n_2-2)!...$ separate color-orderings of the same cut, where
$n_1$, $n_2$... are the multiplicities of the amplitudes making up the
cut. As an example, consider the ``bowtie'' cut topology -- one of the
factorization channels of the iterated unitarity cut -- that has two
topologically distinct color-ordered expansions
\begin{equation}
  \label{eq:kt-color-ord}
   \NMCutC =
  \begin{cases}
     s\text{-}t \text{ channel}
    &
      \frac{1}{\textcolor{ucp-color}{s_{12}}}
      n \left[\scalebox{0.65}{\KTBTContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]\\
    t\text{-}u \text{ channel}
    & 
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
      n\left[
      \scalebox{0.65}{\KTDBContrib{1}{3}{4}{2}}
      \right]
  \end{cases}
\end{equation}
which serve as a complete basis for the full cut.

@@

\subsection{One-loop color-dual integrands}\label{oneLoopCK}
Our first application of this theory for color-dual construction at
loop level is for self-dual Yang-Mills (SDYM). At tree-level, the
off-shell cubic vertices can be written in terms of light cone
coordinates \cite{Monteiro2011pc}:
\begin{equation}
X(p,k) = p_u k_w-p_w k_u
\end{equation}
At loop-level, this construction needs to be analytically continued to
general dimension in order to apply dimensional regularization at
one-loop. This can be done using the cubic semi-YM vertices of the
previous section:
\begin{equation} \label{DdimSDYM}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= \mathcal{X}(k_1,k_2) = (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3) -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
Plugging in on-shell all-plus helicity states in light-cone gauge will
yield precisely the 4D SDYM vertex, up to an unphysical phase:
  \be
  \mathcal{X}(k^+_1,k^+_2) = \frac{\langle 12\rangle^3}{ \langle23\rangle \langle 31\rangle} \sim \langle 12\rangle = k_{1,u} k_{2,w}-k_{1,w} k_{2,u}
 \ee
where $\mathcal{X}(k^+_1,k^+_2) = A^{\text{YM}}(1^+,2^+,3^-)$ in light cone gauge, and in the second step applied momentum conservation to redefinition of the spinor bracket, $\langle 12\rangle \rightarrow X(p,k)$ defined in \cref{Conventions}. Of course, the form of \cref{DdimSDYM} has the advantage of
permitting a $D$-dimensional construction of the one-loop
integrand. Below we provide a four-point example: \nhp{change ordering on legs}
\begin{equation}
\simplebox = g_{\text{YM}}^4 \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\mathcal{X}(k_3,l_3)\mathcal{X}(k_4,l_4)\rangle
\end{equation}
where $l_i = l+k_1+k_2+\cdots+ k_i$ and the bracket $\langle \,\cdots\rangle$ indicates that we have applied the gauge fixed state projector, $\sum {\varepsilon^\mu_{(+l_i)}\varepsilon^\nu_{(-l_i)}}=\eta^{\mu\nu}$ on all internal polarizations. In general, the $n$-gon diagram would be,
\begin{equation}
\mathcal{N}^{\text{SDYM}}_{n\text{-gon}}= \nGonVector = g_{\text{YM}}^n \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\cdots \mathcal{X}(k_n,l_n) \rangle
\end{equation}
Due to the state-sum of internal loop factors,
$\sum {\varepsilon_{(+l)}\varepsilon_{(-l)}}\sim D$, the integrand above
depends explicitly on the spacetime dimension, $D$. By taking a
derivative\footnote{Another way to understand this derivative is that
  it selects the large $D$ behavior of the integrand.}, we can recover
the integrand needed for all-plus one-loop amplitudes with an internal
scalar:
\begin{equation}
\partial_D \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}=\nGonScalar  = (2 g_{\text{YM}})^n (l_1 \varepsilon_1)(l_2 \varepsilon_2)\cdots (l_n \varepsilon_n)
\end{equation}
This integrand numerator with internal scalar loop is precisely what
one would obtain from the Feynman rules of $YZ$-theory, absent the $\bar{Z}Z$
internal vector loop. For more background, we refer the reader to
\cref{sec:CKLagrangians}. When plugging on the on-shell states of $YZ$-theory, we thus obtain the following expression for the NLSM one-loop $n$-gon numerator:
\begin{equation}
 \mathcal{N}^{\text{NLSM}}_{n\text{-gon}}=\left[\partial_D  \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}\right]^{\epsilon \rightarrow k}=\nGonScalar =  f_\pi^{-n} [12][23]\cdots [n1]
\end{equation}
Where we have define the antisymmetric kinematic variable, $[ij] = l_i^2 - l_j^2$, since when plugging on longitudinal modes, the $n$-gon can be re-expressed completely in terms of inverse propagators, $2(k_i \cdot l_i) = l_{i}^2-l^2_{i-1}$. We have verified through 10-point one-loop that this $n$-gon expression is a valid color-dual representation for NLSM.

@@

\section{Two-loop four-point bootstrap}
\label{2loopBoot}

As we have just seen, it is possible to coerce tree numerators into
1-loop numerators at any multiplicity for pions (and related
theories), but that these methods cannot be reapplied to generate
higher-loop numerators.  For pions in particular, the no-go statement
is only for \emph{one particular representation of the theory} so does
not completely preclude the existence of a two-loop color-dual
representation.  We thus turn to the color-dual bootstrap method
described in \cref{sec:bootstrap} to construct a color-dual
representation of two-loop four-point NLSM.  Because of the similarity
of the problem setup, we will also use the opportunity to revisit the
construction of a color-dual representation of pure YM, extending the
search space beyond what was covered in Ref.~\cite{Bern:2015ooa}.

Both representations share the same cubic graph basis, and thus their
defining Jacobi relations lead to the same numerator basis.
Ignoring tadpoles and BEL graphs, there are 14 cubic four-point two-loop
graphs (see \cref{fig:MaxCuts}) which are related to each other via 21
Jacobi relations.  As mentioned in \cref{sec:bootstrap},
color-kinematics duality ensures that the numerator of every graph can
be expressed in terms of a basis of the double box and penta-triangle,
\begin{equation}
\label{eq:jacobi-basis}
\NLSMBasisDoubleBox \text{ and } \NLSMBasisPentaTriangle .
\end{equation}
However, the actual numerator dressings and physical properties
differ, so we discuss them separately below.

@@

For a scalar theory like NLSM the numerator only depends on dot
products of momenta.  The basis of momentum invariants for both of the
graphs can be taken to be
\begin{equation}
\label{eq:scalar-basis}
N^{\text{NLSM}}\left[
  \scalebox{0.75}{\NLSMBasisDoubleBox} \right]
\text{ and }
N^{\text{NLSM}}\left[\scalebox{0.75}{\NLSMBasisPentaTriangle} \right]\in
\left\{\begin{array}{cccc}
    k_{13},& k_{23},& k_{15},& k_{16}, \\
    k_{25},& k_{26},& k_{35},& k_{36}, \\
    k_{55},& k_{56},& k_{66}
  \end{array}
  \right\},
\end{equation}
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_6 = \ell_2$.
%%%%
% Basis for both diagrams
% {d[p[1], p[3]], d[p[1], p[6]], d[p[1], p[7]], d[p[2], p[3]],  d[p[2], p[6]], d[p[2], p[7]], d[p[3], p[6]], d[p[3], p[7]],  d[p[6], p[6]], d[p[6], p[7]], d[p[7], p[7]]}
%%%%
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram
scales as $k^{12}$.  At this mass dimension there are 8,008
independent dot products of momenta.  With two basis graphs there is a
total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is
  not suited to solving large sets of equations with sixteen thousand
  parameters, so a different method is desirable.  One possibility is
  to use a sprase solver, as often employed for integral reduction.
  However, these solvers typically assume that the matrix has a very
  low density which is not the case for equations coming from NLSM
  which typically have a density of a few percent.  While the
  equations in YM naturally separate into different ``helicity''
  sectors (or $e_i \cdot e_j$ and $e_i \cdot k_j$ sectors in $D$
  dimensions), there is no such separation of sectors in a scalar
  theory where any product of momenta can interfere with any other.
  The NLSM constraint equations were solved using a custom solver,
  \texttt{FiniteFieldSolve}, designed to solve large linear systems of
  arbitrary density.  \texttt{FiniteFieldSolve} will be released
  publicly shortly.\ace{I want to rephrase this slightly, but need to
    think about it a bit more}} Imposing the boundary Jacobi relations
eliminates all but 4,473 of the original free parameters.  Further
imposing graph symmetries reduces the number of free parameters to
1,243.

@@

\begin{figure}[h]
  \begin{center}
	\PhysicalCutOne{}{}{}{} \PhysicalCutTwo{}{}{}{}
  \end{center}
  \caption{The two physical cut topologies shared by NLSM, sGal, BI, and DBIVA}
  \label{fig:emu}
\end{figure}

@@

\ace{Lets tighten up this paragraph a bit, especially the second half}
It is tempting to try to match the pion four-point two-loop integrand
to ZM on the cuts, but given that the cuts only probe the on-shell
four-point amplitude, which is either subtle or vanishes for ZM, it is
better to compare the off-shell numerators directly.  As local
functions, the numerators never suffer from any of the subtleties of
on-shell 2D kinematics.  Since the physical cuts are difficult to
compare, ZM theory is being used here simply as a tool to eliminate
generalized gauge freedom; ZM theory has the right power counting and
kinematic structure in 2D to compare against.  Since ZM theory is only
being used for the kinematical structure of its numerators, it is
natural to introduce a proportionality constant $z$ between the NLSM
and ZM numerators $n_\text{NLSM} \vert_{2D} = z ~ n_\text{ZM}$.  The
parameter $z$, which would normally be fixed on cuts, effectively
encodes the discrepancy between the coupling constants of the two
theories.\footnote{Floating the proportionality parameter $z$ is also
  motivated because ZM is believed to be quantum mechanically
  different from NLSM even though they are dual classically in two
  dimensions \cite{Nappi:1979ig}.  When the four-point
  pion amplitude is normalized to $-k_1\cdot k_2$ and the ZM vertex is
  normalized to $k_1^\mu \varepsilon_{\mu\nu}k_2^\nu$, the parameter
  $z$ is fixed to $216/565$, which is clearly unrelated to the
  difference in coupling constants.}  This introduces an extra degree of
freedom into the problem.  Mechanically, the pion numerator is matched
to ZM by first taking every possible assignment of right and left
movers for the external particles and then restricting the pion
numerator to 2D.  The loop momenta are restricted to 2D but left
off-shell.  After performing the 2D matching, there are only 365
parameters of generalized gauge freedom.

The loop momentum structure of ZM theory provides one final hint for
simplifying the pion numerators.  A generic term in the ZM numerator
(of either basis graph) looks schematically like
\begin{equation}
\label{eq:ZMLoopPowCount}
\ell_1^m \ell_2^n k^{12-(m+n)} \text{ where } 5 \leq m+n \leq 8,
\end{equation}
whereas a generic set of local, cubic Feynman rules could have
produced terms with up to $m+n=12$ powers of loop momenta.  When the
pion numerator is forced to have the loop power counting structure in
\cref{eq:ZMLoopPowCount} of ZM theory, the number of generalized gauge
freedom parameters reduces to 58.  The pion numerator appearing in the
ancillary files makes exactly this choice.

@@

Given the close ties between pion and gluon scattering for trees and
at one loop, the existence of the two-loop pion numerator prompts us
to investigate the most general local numerator for YM, without any of
the loop power counting assumptions of Ref.~\cite{Bern:2015ooa}.  We
follow the same general procedure as in \cref{sec:pions}, again
identifying the double-box and penta-triangle as the
Jacobi basis graphs and building a the most general ansatz for each of
their numerators compatible with the assumptions in
\cref{sec:bootstrap}.  Because we are now considering pure
Yang-Mills, each monomial in the local ansatz must consist of five
Lorentz scalar dot product s instead of the six for NLSM, and each
monomial must be linear in each of the four external gluon
polarizations.  Thus, the numerator ansatz for both diagrams will be
built from terms of the form:
\begin{equation}
  \mathcal{N}^{\text{YM}}\left[\scalebox{0.75}{\NLSMBasisDoubleBox} \right]
  \text{ and }
  \mathcal{N}^{\text{YM}}\left[\scalebox{0.75}{\NLSMBasisPentaTriangle} \right]
  \in \text{span}\left\{(\varepsilon_i \varepsilon_j) , (k_i \varepsilon_j), (k_ik_j)\right\}
  \label{eq:ym-basis}
\end{equation}
where the $k_i$ take the same definition as described near
\cref{eq:scalar-basis}.  In our labeling of the two loop Jacobi-basis
diagrams, shown in \cref{eq:jacobi-basis}, we find that the power
counting bound proposed by Bern, Davies, and Nohle leads to an initial
ansatz with 10,000 terms in the double-box, and 9849 terms for the
penta-triangle.  Without power counting restrictions, both the
double-box and the penta-triangle ansatz have an ansatz with 10,010
terms each.  Imposing diagram automorphisms and maximal-cut gauge
invariance on the two basis diagrams, we reduce the number of terms to
XX for the double-box, and YY for the penta-triangle.  In
Ref.~\cite{Bern:2015ooa}, the unitarity matching constraints were
imposed using a spanning set of minimal physical cuts, shown in
\cref{fig:ym-spanning}.  Instead of also directly going to the
spanning cuts, we work within the framework of the Method of Maximal
Cuts \cite{Bern:2007ct} in order to identify the simplest cut that is in
tension with the kinematic Jacobi relations.

>>>>

In color-charged theories, both sides of the cut can be further
decomposed according to the color algebra.  For theories charged under
the adjoint of $SU(N)$ via $f^{abc}$-dressed amplitudes, it is
convenient to project onto a specific element of the Del
Duca-Dixon-Maltoni (DDM) color basis \cite{DixonMaltoni} by only
inserting color-ordered amplitudes, a la the right hand side of
\cref{eq:ddm-basis} in the product of trees, and restricting the set
of compatible diagrams to those whose color factors reduce to the
appropriate DDM element when summing over \emph{only the uncut color
  contractions}.  In this situation, the color becomes an overall
factor on the entire cut equality and thus can be ignored.  Cuts that
are organized in this manner are known as ``color-ordered'' cuts, in
the same sense as color-ordered tree amplitudes.  Since the DDM basis
(and dual Kleiss-Kuijf amplitude basis \cite{Kleiss:1988ne}) have
$(n-2)!$ elements, one generally needs to evaluate
$(n_1-2)! (n_2-2)!...$ separate color-orderings of the same cut, where
$n_1$, $n_2$... are the multiplicities of the amplitudes making up the
cut. As an example, consider the ``bowtie'' cut topology -- one of the
factorization channels of the iterated unitarity cut -- which has two
topologically distinct color-ordered expansions
\begin{equation}
  \label{eq:kt-color-ord}
   \NMCutC =
  \begin{cases}
     s\text{-}t \text{ channel}
    &
      \frac{1}{\textcolor{ucp-color}{s_{12}}}
      n \left[\scalebox{0.65}{\KTBTContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]\\
    t\text{-}u \text{ channel}
    & 
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
      n\left[
      \scalebox{0.65}{\KTDBContrib{1}{3}{4}{2}}
      \right]
  \end{cases}
\end{equation}
which serve as a complete basis for the full cut.

@@

\subsection{One-loop color-dual integrands}\label{oneLoopCK}
Our first application of this theory for color-dual construction at
loop level is for self-dual Yang-Mills (SDYM). At tree-level, the
off-shell cubic vertices can be written in terms of light cone
coordinates \cite{Monteiro2011pc}:
\begin{equation}
X(p,k) = p_u k_w-p_w k_u
\end{equation}
At loop-level, this construction needs to be analytically continued to
general dimension in order to apply dimensional regularization at
one-loop. This can be done using the cubic semi-YM vertices of the
previous section:
\begin{equation} \label{DdimSDYM}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= \mathcal{X}(k_1,k_2) = (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3) -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)
\end{equation}
Plugging in on-shell all-plus helicity states in light-cone gauge will
yield precisely the 4D SDYM vertex, up to an unphysical phase \ace{Can
  we provide an example?} \nhp{yes, provided}:
  \be
  \mathcal{X}(k^+_1,k^+_2) = \frac{\langle 12\rangle^3}{ \langle23\rangle \langle 31\rangle} \sim \langle 12\rangle = k_{1,u} k_{2,w}-k_{1,w} k_{2,u}
 \ee
where $\mathcal{X}(k^+_1,k^+_2) = A^{\text{YM}}(1^+,2^+,3^-)$ in light cone gauge, and in the second step applied momentum conservation to redefinition of the spinor bracket, $\langle 12\rangle \rightarrow X(p,k)$ defined in \cref{Conventions}. Of course, the form of \cref{DdimSDYM} has the advantage of
permitting a $D$-dimensional construction of the one-loop
integrand. Below we provide a four-point example: \nhp{change ordering on legs}
\begin{equation}
\simplebox = g_{\text{YM}}^4 \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\mathcal{X}(k_3,l_3)\mathcal{X}(k_4,l_4)\rangle
\end{equation}
where $l_i = l+k_1+k_2+\cdots+ k_i$ and the bracket $\langle \,\cdots\rangle$ indicates that we have applied the gauge fixed state projector, $\sum {\varepsilon^\mu_{(+l_i)}\varepsilon^\nu_{(-l_i)}}=\eta^{\mu\nu}$ on all internal polarizations. In general, the $n$-gon diagram would be,
\begin{equation}
\mathcal{N}^{\text{SDYM}}_{n\text{-gon}}= \nGonVector = g_{\text{YM}}^n \langle \mathcal{X}(k_1,l_1)\mathcal{X}(k_2,l_2)\cdots \mathcal{X}(k_n,l_n) \rangle
\end{equation}
Due to the state-sum of internal loop factors,
$\sum {\varepsilon_{(+l)}\varepsilon_{(-l)}}\sim D$, the integrand above
depends explicitly on the spacetime dimension, $D$. By taking a
derivative\footnote{Another way to understand this derivative is that
  it selects the large $D$ behavior of the integrand.}, we can recover
the integrand needed for all-plus one-loop amplitudes with an internal
scalar:
\begin{equation}
\partial_D \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}=\nGonScalar  = (2 g_{\text{YM}})^n (l_1 \varepsilon_1)(l_2 \varepsilon_2)\cdots (l_n \varepsilon_n)
\end{equation}
This integrand numerator with internal scalar loop is precisely what
one would obtain from the Feynman rules of $YZ$-theory, absent the $\bar{Z}Z$
internal vector loop. For more background, we refer the reader to
\cref{sec:CKLagrangians}. When plugging on the on-shell states of $YZ$-theory, we thus obtain the following expression for the NLSM one-loop $n$-gon numerator:
\begin{equation}
 \mathcal{N}^{\text{NLSM}}_{n\text{-gon}}=\left[\partial_D  \mathcal{N}^{\text{SDYM}}_{n\text{-gon}}\right]^{\epsilon \rightarrow k}=\nGonScalar =  f_\pi^{-n} [12][23]\cdots [n1]
\end{equation}
Where we have define the antisymmetric kinematic variable, $[ij] = l_i^2 - l_j^2$, since when plugging on longitudinal modes, the $n$-gon can be re-expressed completely in terms of inverse propagators, $2(k_i \cdot l_i) = l_{i}^2-l^2_{i-1}$. We have verified through 10-point one-loop that this $n$-gon expression is a valid color-dual representation for NLSM.

@@

\section{Two-loop bootstrap}
\label{2loopBoot}
\ace{Lets bulk this section intro up a bit} \nhp{@Alex will bulk this up}

We now employ the color-dual bootstrap method described in
\cref{sec:bootstrap}. We will start with an ansatz constructed from
$D$-dimensionally Lorentz covariant kinematics. To fix remaining
freedom in the NLSM bootstrap, we will match down to ZM theory in
$D=2$. This is merely an aesthetic choice, so the final results of
both the NLSM and Yang-Mills bootstraps are dimensionally agnostic.

@@

As we have seen, it is possible to coerce tree numerators into 1-loop
numerators at any multiplicity for pions.  However, these methods
cannot be reapplied to generate higher-loop numerators, which begs the
question ``is it possible to find globally color-dual numerators
beyond 1-loop?''  We will tackle this question for four-point two-loop
scattering in NLSM before moving on to the same process in YM.

Ignoring tadpoles and BEL graphs, there are 14 cubic four-point two-loop
graphs (see \cref{fig:MaxCuts}) which are related to each other via 21
Jacobi relations.  As mentioned in \cref{sec:bootstrap},
color-kinematics duality ensures that the numerator of every graph can
be expressed in terms of a basis of the double box and penta-triangle,
\begin{equation}
\label{eq:jacobi-basis}
\NLSMBasisDoubleBox \text{ and } \NLSMBasisPentaTriangle .
\end{equation}
For a scalar theory like NLSM the numerator only depends on dot
products of momenta.  The basis of momentum invariants for both of
these graphs can be taken to be
\begin{equation}
\label{eq:scalar-basis}
\{ k_{13}, k_{23}, k_{15}, k_{16}, k_{25}, k_{26}, k_{35}, k_{36}, k_{55}, k_{56}, k_{66} \},
\end{equation}
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_6 = \ell_2$.
%%%%
% Basis for both diagrams
% {d[p[1], p[3]], d[p[1], p[6]], d[p[1], p[7]], d[p[2], p[3]],  d[p[2], p[6]], d[p[2], p[7]], d[p[3], p[6]], d[p[3], p[7]],  d[p[6], p[6]], d[p[6], p[7]], d[p[7], p[7]]}
%%%%
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram
scales as $k^{12}$.  At this mass dimension there are 8,008
independent dot products of momenta.  With two basis graphs there is a
total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is
  not suited to equations with sixteen thousand parameters so a
  different method is desirable.  One possibility is to use a solver
  designed for integral reduction.  However, these solvers typically
  assume that the matrix has a very low density which is not the case
  for equations coming from NLSM which typically have a density of a
  few percent.  While the equations in YM naturally separate into
  different ``helicity'' sectors (or $e_i \cdot e_j$ and
  $e_i \cdot k_j$ sectors in $D$ dimensions), there is no such
  separation of sectors in a scalar theory where any product of
  momenta can interfere with any other.  The NLSM constraint equations
  were solved using a custom solver, \texttt{FiniteFieldSolve},
  designed to solve large linear systems of arbitrary density.
  \texttt{FiniteFieldSolve} will be released publicly shortly.\ace{I
    want to rephrase this slightly, but need to think about it a bit
    more}} Imposing the boundary Jacobi relations eliminates all but
4,473 of the original free parameters.  Further imposing graph
symmetries reduces the number of free parameters to 1,243.

\begin{figure}
  \begin{center}
	\PhysicalCutOne{}{}{}{} \PhysicalCutTwo{}{}{}{}
  \end{center}
  \caption{The two physical cut topologies shared by NLSM, sGal, BI, and DBIVA}
  \label{fig:emu}
\end{figure}

@@

@@

\ace{Lets tighten up this paragraph a bit, especially the second half}
It is tempting to try to match the pion four-point two-loop integrand to ZM
on the cuts, but given that the cuts only probe the on-shell four-point
amplitude, which is either subtle or vanishes for ZM, it is better to
compare the off-shell numerators directly.  As local functions, the
numerators never suffer from any of the subtleties of on-shell 2D
kinematics.  Since the physical cuts are difficult to compare, ZM
theory is being used here simply as a tool to eliminate generalized
gauge freedom; ZM theory has the right power counting and kinematic
structure in 2D to compare against.  Since ZM theory is only being
used for the kinematical structure of its numerators, it is natural to
introduce a proportionality constant $z$ between the NLSM and ZM
numerators $n_\text{NLSM} \vert_{2D} = z ~ n_\text{ZM}$.  The
parameter $z$, which would normally be fixed on cuts, effectively
encodes the discrepancy between the coupling constants of the two
theories.\footnote{Floating the proportionality parameter $z$ is also
  motivated because ZM is believed to be quantum mechanically
  different from NLSM even though they are dual classically in two
  dimensions \cite{Nappi:1979ig}.}  This introduces an extra degree of
freedom into the problem.  Mechanically, the pion numerator is matched
to ZM by first taking every possible assignment of right and left
movers for the external particles and then restricting the pion
numerator to 2D.  The loop momenta are restricted to 2D but left
off-shell.  After performing the 2D matching, there are only 365
parameters of generalized gauge freedom.

The loop momentum structure of ZM theory provides one final hint for
simplifying the pion numerators.  A generic term in the ZM numerator
(of either basis graph) looks schematically like
\begin{equation}
\label{eq:ZMLoopPowCount}
\ell_1^m \ell_2^n k^{12-(m+n)} \text{ where } 5 \leq m+n \leq 8,
\end{equation}
whereas a generic set of local, cubic Feynman rules could have
produced terms with up to $m+n=12$ powers of loop momenta.  When the
pion numerator is forced to have the loop power counting structure in
\cref{eq:ZMLoopPowCount} of ZM theory, the number of generalized gauge
freedom parameters reduces to 58.  The pion numerator appearing in the
ancillary files makes exactly this choice.\footnote{When the four-point
  pion amplitude is normalized to $-k_1\cdot k_2$ and the ZM vertex is
  normalized to $k_1^\mu \varepsilon_{\mu\nu}k_2^\nu$, the parameter
  $z$ is fixed to $216/565$, which is clearly unrelated to the
  difference in coupling constants.} \ace{Is this actually where we
  want this particular footnote to appear?}

@@

\jm{Rather than quoting the number of parameters in the BDN ansatz, I
  would prefer to just focus on what we did since it's more general.
  I don't want it to sound like we just patched a minor hole in their
  argument when I think you've done something much cooler by
  pinpointing the failure to the bowtie graph. I know you put time and
  effort into analyzing the BDN ansatz but I'm worried that the more
  we talk about it, the more it will overshadow the real success
  here.}

Given the close ties between pion and gluon scattering for trees and
at one loop, the existence of the two-loop pion numerator prompts us
to investigate the most general local numerator for YM, without any of
the loop power counting assumptions of Ref.~\cite{Bern:2015ooa}.  We
follow the same general procedure as in \cref{sec:pions}, again
identifying the double-box (2box) and penta-triangle (pt) as the
Jacobi basis graphs and building a the most general ansatz for each of
their numerators compatible with the assumptions in
\cref{sec:bootstrap}.  Because we are now considering pure
Yang-Mills, each monomial in the local ansatz must consist of five
Lorentz scalar dot product s instead of the six for NLSM, and each
monomial must be linear in each of the four external gluon
polarizations.  Thus, the numerator ansatz for both diagrams will be
built from terms of the form:
\begin{equation}
  \mathcal{N}^{\text{YM}}\left[\NLSMBasisDoubleBox \right]\text{ and }\mathcal{N}^{\text{YM}}\left[\NLSMBasisPentaTriangle \right]\in \text{span}\left\{(\varepsilon_i \varepsilon_j) , (k_i \varepsilon_j), (k_ik_j)\right\}
  \label{eq:ym-basis}
\end{equation}
where the $k_i$ take the same definition as described near
\cref{eq:scalar-basis}.  In our labeling of the two loop Jacobi-basis
diagrams, shown in \cref{eq:jacobi-basis}, we find that the power
counting bound proposed by Bern, Davies, and Nohle leads to an initial
ansatz with 10,000 terms in the double-box, and 9849 terms for the
penta-triangle.  Without power counting restrictions, both the
double-box and the penta-triangle ansatz have an ansatz with 10,010
terms each.  Imposing diagram automorphisms and maximal-cut gauge
invariance on the two basis diagrams, we reduce the number of terms to
XX for the double-box, and YY for the penta-triangle.  In
Ref.~\cite{Bern:2015ooa}, the unitarity matching constraints were
imposed using a spanning set of minimal physical cuts, shown in
\cref{fig:ym-spanning}.  Instead of also directly going to the
spanning cuts, we work within the framework of the Method of Maximal
Cuts \cite{Bern:2007ct} in order to identify the simplest cut that is in
tension with the kinematic Jacobi relations.
